PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

### 2024-03-11 docs/_tex/index.tex
Bede's comments

\begin{abstract}
Coastal areas host seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges and reduce the
issue of mlixed pixel, drone images at a very high spatial resolution
offer a promising solution. This study focuses on using drone
acquisitions with a ten spectral bands sensor mirroring those of
Sentinel-2, for mapping intertidal macrophytes and effectively
discriminating between seagrass and green macroalgae. Nine drone flights
were conducted at two different altitudes (12m and 120m) across diverse
European habitats in France and Portugal. Low altitude flights were used
to train a Deep Learning classifier based on Neural Networks to
discrimintate among five taxonomic classes of intertidal vegetation:
Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae), Phaeophyceae
(Brown algae), Rhodophyceae (Red macroalgae) and Bacillariophyceae
(Diatoms). Drone mapping demonstrated an overall accuracy of 94\% across
all the sites and images, covering a total area of 467 000 m². The model
exhibited an accuracy of 96.4\% in identifying seagrass. This work
showed that although they share the same pigmentary composition,
seagrass and green algae can be discriminated using ten bands
multispectral sensor. The algorithm should now be adapted for Sentinel-2
to see if this discrimination is still operational at a coarser spatial
resolution.
\end{abstract}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between land and ocean \citep{unsworth2022}. Seagrass meadows provide a
myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp{duffy2019}}. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scale, many
coastal waters worldwide are still subjected to strong eutrophication
processes \citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}.
Coastal eutrophication has been ass ociated to anomalous accumulation of
green macroalgae, the so-called green tides
(\citep{devlin2023nutrients}). Green tides produce shade and suffocation
over seagrass individuals, thus threatening the health of seagrass
ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Seagrass monitoring is all the more important as it makes possible to
identify which pressure should be addressed in priority to facilitate
effective mitigation actions. Traditionally, indicators of seagrass
status have been quantified using \emph{in situ} measurements. The
acquisition of field data in intertidal zones is however notoriously
challenging. Intertidal seagrass meadows are only partially exposed
during low tide and can be situated in difficult-to-reach mudflats,
potentially leading to inaccurate and limited estimations with
conventional sampling techniques \citep{nijland2019}. Satellite
observations have been proven effective in complementing in situ
sampling, allowing for the near real-time and consistent retrieval of
seagrass EOVs and EBVs over extensive meadows. \citetext{\citealp[
]{Zoffoli2021}; \citealp[ ]{xu2021}; \citealp[
]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Mission with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. As a result,
their respective spectral signatures can be considered similar by a
non-expert observer \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance spectra, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. The application of this approach to satellite remote sensing
remains however to be validated. Moreover patches of green algae can
develop at small spatial scales that are not observable using
non-commercial satellite imagery \citep{tuya2013}, especially during the
initial stage of the event.

Drones can potentially fill the gaps left by satellite remote sensing
and in situ measurements, due to their ability to provide
spatially-explicit observations at very high spatial resolutions (pixel
size from cm to mm) and still capturing data at multi-spectral
resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most study cases were limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone remote sensing
to map intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae. Ten drone flights were performed over
soft-bottom intertidal areas along two European countries (France and
Portugal), covering a wide range of habitats, from monospecific seagrass
meadows to meadows mixed with green, or red macroalgae. A deep learning
algorithm was trained and validated for macrophyte discrimination,
emphasizing applicability across diverse sites without a loss of
prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : one is located in one of these
island (Arz island); the other one is located in the southern part of
the gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. Bougneuf bay hosts a large
intertidal seagrass meadow of about 6 km². Within this meadow, the sites
observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contain a monospecific bed of \emph{Zostera noltei} with very little
mixing with other macrophytes. Three sites have been surveyed in the Ria
de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this
lagoon is about 80 km² with many narrow channels, large salt marshes and
many mudflats that uncovers at low tide. It is connected to the open sea
through a single channel, with a tidal lag between the North and the
South of the lagoon. The southernmost site (Gafanha) is a mudflat
located in an affluent of the lagoon whereas the two other sites are
situated in the middle of the lagoon and are only accessible by boat
(Figure~\ref{fig-map} C). These Portuguese sites are characterized by a
more diverse intertidal vegetation, where patches of seagrass could
intermingle with red, brown and green macroalgae.

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourngeuf Bay (Two sites),
C: Ria de Aveiro Lagoon (Three sites). Green represent intertidal
areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in ten spectral bands, from 444 nm
(blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. A structure-from-motion photogrammetry
software (Agisoft Metashape) was used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking was the same for every flight. First, tying key points
were detected inside of each image and between overlapping images in
order to obtain a sparse point cloud. This cloud was cleaned using
reprojection accuracy metric in order to remove noisy points. A dense
point cloud was then produced using a structure from motion algorithm. A
surface interpolation of this dense point cloud was made to obtain a
digital surface model (DSM), used to reconstruct the multispectral
ortho-image. Across all sites, flights were made at two different
altitudes : 12 m or/and 120 m. Low altitude drone flights produce
ortho-images with a very high spatial resolution (8 mm per pixel),
making it efficient to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allow to cover large
areas and produced images with a pixel size of 80 mm
(Table~\ref{tbl-flights}).

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² assessing the presence or absence of five key taxonomic
classes of intertidal vegetation : Bacillariophyceae (Unicellular
benthic diatoms forming large biofilms at the sediment surface during
low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (seagrass),
Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae)
(Figure~\ref{fig-vegetation}). Pictures of each quadrat were uploaded
online on the Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}, a public open portal to store and share biodiversity
data. Each photograph was also processed to estimate the percent cover
of each type of vegetation using an image processing software (imageJ).
For all quadrats, the hyperspectral reflectance signatures of each
vegetation class was measured using an ASD FieldSpec HandHeld 2
spectroradiometer, which acquires reflectance between 325 and 1075 nm,
with 1 nm of spectral resolution. Hyperspectral signatures from quadrats
serve dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretation. The
workflow developped in this study is presented in
Figure~\ref{fig-workflow}.

@@

The spectral similarities of the reflectance signatures between
intertidal green macrophytes (Magnoliopsida and Chlorophyceae) make
their discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was developed, trained, validated, and
applied to each drone flight (Figure~\ref{fig-workflow}).

@@

The model was applied to all flights at both 12 and 120 m of altitude.
In situ information on georeferenced class type and percent cover
collected during each flight was used to assess the model accuracy.
These images were used to construct a validation dataset indicating the
presence or absence of each class. Additionally to the quadrat-based
validation dataset, polygons of each class were photo interpreted in
order to increase the number of pixels of the validation dataset. A
confusion matrix, along with precision metrics such as global accuracy,
sensitivity, specificity, and Kappa coefficient, was generated for each
sites. All validation matrices were then merged to create a unique
matrix . Altogether, a total of 536,000 pixels was used to validate the
model, thus providing a geographically robust validation dataset.

@@

A total of nine prediction maps corresponding to the nine drone flights
were obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the five
vegetation classes on which the model was trained, four were present on
this site, with Chlorophyceae and Rhodophyceae mixed with the seagrass
meadow. There was also Bacillariophyceae forming biofilms on parts of
sediment surface. Although the seagrass bed was solely composed of
\emph{Zostera noltei}, various colors of this could be observed from
dark green (corresponding to healthy leaves) to whitish/brown (when
leaves were discolored having lost their pigmentation). Regardless of
color, the meadow was predicted as Magnoliopsida by the model.

@@

\caption{\label{fig-GafLow}RGB orthomosaic (Left) and Prediction (Right)
of the low altitude flight of Gafanha, Portugal. The total extent of
this flight is 3000m² with a resolution of 8 mm per pixel. Background
colors indicate intertidal area (Light Green) and land area (Light
Grey). The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel
size.}

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
masked in the prediction map due to sunglint and misclassification
caused by the turbid shallow water. Most of the intertidal area was
classified as Magnoliopsida by the model, including seagrass patches
with discolored leaves. Only a few pixels were classified as
Chlorophyceae at this scale. Furthermore, the area that was classified
as Bacillariophyceae in the low-altitude flight remained mostly
classified as such in the high-altitude flight, though some pixels were
classified as Magnoliopsida. Patches of Rhodophyceae were correctly
classified. In the nothern part of the scene, near the land limit,
patches of the schorre angiosperm \emph{Spartina sp.} were
misclassified, either as Magnoliopsida or as Phaeophyceae.

@@

\caption{\label{fig-GafHigh}RGB orthomosaic (Left) and Prediction
(Right) of the high altitude flight of Gafanha, Portugal. The total
extent of this flight is about 1 km² with a resolution of 80 mm per
pixel. Background colors indicate intertidal area (Light Green), land
area (Light Grey) and water (Light Blue). The red triangle shows the
extent of the low altitude flight of Gafanha presented in
Figure~\ref{fig-GafLow}. The zoom covers an area equivalent to a
10-meter Sentinel-2 pixel size.}

@@

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight is about 1.5 km² with a resolution of 80
mm per pixel. Background colors indicates intertidal area (Light Green),
land area (Light Grey) and water (Light Blue). Each cover an area
equivalent to a 10-meter Sentinel-2 pixel size.}

@@

\caption{\label{fig-Dike}RGB orthomosaic (Top) and Prediction (Bottom)
of the Northern part of Noirmoutier Island, France. The total extent of
this flight is about 28 000 m² with a resolution of 80 mm per pixel.
Background colors indicate intertidal area (Light Green) and land area
(Light Grey). The zoom covers an area equivalent to a 10-meter
Sentinel-2 pixel size.}

@@

A total of 536,000 pixels was used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Kerdrean (5557 pixels) and Marinha Lanzarote (159713 pixels),
respectively. Model global accuracy was 94.26\% with a Kappa coefficient
of 0.92 (Figure~\ref{fig-Validation}). The least performing site was
Gafanha High (global accuracy of 75.45\%) whereas Mataducos was the site
with the most accurate prediction (global accuracy of 98.05\%). Overall,
the classes Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae were
correctly classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae was the least performing class (accuracy
of 0.72 ) mainly due to the confusion with Magnoliopsida and Sediment.

@@

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each taxonomic class. The bigger the slice, the more
important is the variable to predict accurately this class. The right
plot shows the drone raw and standardised reflectance spectra of each
class. Each slice represent the VIP of both raw and standardised
reflectance combined.}

@@

\caption{\label{fig-upscaling}Kernel density plot showing the proportion
of pixel well classified based on the percent cover of the class in high
altitude flight pixels of Gafanha, Portugal. Each subplot shows all the
pixels of the same classes on the hight altitude flight. Percent cover
of classes was retrieved using the result of the classification of the
low altitude flight of Gafanha, Portugal. The vertical dashed line shows
the 0.85 probability of the model. Everything on the right of this line
has a probability higher then 0.85 and everything on the left of this
line has a lower probability.}

>>>>

\begin{abstract}
Coastal areas support seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges and reduce the
issue of mlixed pixel, drone images at a very high spatial resolution
offer a promising solution. This study focuses on using drone
acquisitions with a ten spectral band sensor mirroring those of
Sentinel-2, for mapping intertidal macrophytes and effectively
discriminating between seagrass and green macroalgae. Nine drone flights
were conducted at two different altitudes (12m and 120m) across
heterogeneous European habitats in France and Portugal. Low altitude
flights were used to train a Deep Learning classifier based on Neural
Networks to discrimintate among five taxonomic classes of intertidal
vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae),
Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae) and
Bacillariophyceae (Diatoms). Classification of drone imagery resulted in
an overall accuracy of 94\% across all the sites and images, covering a
total area of 467 000 m². The model exhibited an accuracy of 96.4\% in
identifying seagrass. This work showed that although they share the same
pigmentary composition, seagrass and green algae can be discriminated
using ten bands multispectral sensor. The algorithm should now be
adapted for Sentinel-2 to see if this discrimination is still
operational at a coarser spatial resolution.
\end{abstract}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp[ ]{duffy2019}; \citealp[
]{rasheed2011long}; \citealp{chefaoui2018dramatic}}. While improvements
in water quality have been recently reported in Europe, allowing an
overall recovery of seagrass ecosystems at local scale, many coastal
waters worldwide are still subjected to strong eutrophication processes
\citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}. Coastal
eutrophication has been ass ociated to excessive accumulation of green
macroalgae, so-called green tides (\citep{devlin2023nutrients}). Green
tides produce shade and suffocation over seagrass individuals, thus
threatening the health of seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. The acquisition of field data in intertidal
zones is however notoriously challenging. Intertidal seagrass meadows
are only partially exposed during low tide and can be situated in
difficult-to-reach mudflats, potentially leading to inaccurate and
limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing in situ sampling, allowing for the near real-time and
consistent retrieval of seagrass EOVs and EBVs over extensive meadows.
\citetext{\citealp[ ]{Zoffoli2021}; \citealp[ ]{xu2021}; \citealp[
]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Missions with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. Therefore, to
someone not specialized in the field, their spectral signatures may
appear to be alike \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance spectra, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. However the application of this approach to satellite RS
remains to be validated. Moreover patches of green algae can develop at
small spatial scales that are not observable using non-commercial
satellite imagery \citep{tuya2013}, especially during the initial stage
of th

Drones can potentially fill the gaps left by satellite RS and in situ
measurements, due to their ability to provide spatially-explicit
observations at very high spatial resolutions (pixel size from cm to mm)
while capturing data at multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most case studies are limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone RS to map
intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae (Seagrass and Green Algae respectively).
Ten drone flights were performed over soft-bottom intertidal areas along
two European countries (France and Portugal), covering a wide range of
habitats, from monospecific seagrass meadows to meadows mixed with
green, or red macroalgae. A deep learning algorithm was trained and
validated for macrophyte discrimination, emphasizing applicability
across diverse sites without a loss of prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in the Gulf of Morbihan (Figure~\ref{fig-map} A), France
(47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only
connected to the sea through a 900m wide channel. A total of 53 small
islands are scattered across the gulf leading to 250 km of shorelines.
Patchy seagrass meadows can be found on a lot of these islands. Within
the Gulf of Morbihan we chose two sites : one located on one of the
gulf's islands (Arz island) and the other located in the southern part
of the gulf (Duer). Two other sites were located in Bourgneuf Bay,
France (46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed
macrotidal bay, protected from waves by Noirmoutier Island. Bougneuf bay
hosts a large intertidal seagrass meadow of about 6 km². Within this
meadow, the sites observed by drones (L'Epine and Barbatre,
Figure~\ref{fig-map} B) contained a monospecific bed of
\emph{Nanozostera noltei} (dwarf eelgrass) with very little mixing with
other macrophytes. Three sites were surveyed in the Ria de Aveiro Lagoon
in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is about 80
km² with many narrow channels, large salt marshes and many mudflats that
uncovers at low tide. It is connected to the open sea through a single
channel, with a tidal lag between the North and the South of the lagoon.
The southernmost site (Gafanha) was a mudflat located in a tributary
river of the lagoon whereas the two other sites were situated in the
middle of the lagoon and only accessible by boat (Figure~\ref{fig-map}
C). These Portuguese sites are characterized by a more diverse
intertidal vegetation, where patches of seagrass intermingle with red,
brown and green macroalgae.

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourngeuf Bay (Two sites),
C: Ria de Aveiro Coastal Lagoon (Three sites). Green represent
intertidal areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands, from 444
nm (blue) to 840 nm (near infrared, NIR). To ensure consistent lighting
conditions across flight paths, the drone's trajectory was aligned to
maintain a solar azimuth angle of 90 degrees. An overlap of 70\% and
80\% (side and front respectively) between each image was set for each
flight. A downwelling light sensor (DLS2) was used to acquire irradiance
data concomitantly with the camera measurements. Raw data were
calibrated in reflectance using a calibration panel reflective at
\textasciitilde50\% provided by the manufacturer. A
structure-from-motion photogrammetry software (Agisoft Metashape) was
used to process images to obtain multispectral orthomosaics of each
flight. The workflow for orthomosaicking was the same for every flight.
First, tying key points were detected inside of each image and between
overlapping images in order to obtain a sparse point cloud. This cloud
was cleaned using reprojection accuracy metric in order to remove noisy
points. A dense point cloud was then produced using a structure from
motion algorithm. A surface interpolation of this dense point cloud was
made to obtain a digital surface model (DSM), used to reconstruct the
multispectral ortho-image. Across all sites, flights were made at two
different altitudes : 12 m or/and 120 m. Low altitude drone flights
produce ortho-images with a very high spatial resolution (8 mm per
pixel), making it efficient to visually distinguish between the various
types of vegetation. High altitude fights on the other hand allow to
cover large areas and produced images with a pixel size of 80 mm
(Table~\ref{tbl-flights}).

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² assessing the presence or absence of five key taxonomic
classes of intertidal vegetation : Bacillariophyceae (Unicellular
benthic diatoms forming biofilms at the sediment surface during low
tide), Phaeophyceae (brown macroalgae), Magnoliopsida (seagrass),
Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae)
(Figure~\ref{fig-vegetation}). Pictures of each quadrat were uploaded
online to the Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}, a public open portal to store and share biodiversity
data. Each photograph was also processed to estimate the percent cover
of each type of vegetation using an image processing software (imageJ).
For all quadrats, the hyperspectral reflectance signatures of each
vegetation class was measured using an ASD FieldSpec HandHeld 2
spectroradiometer, which acquires reflectance between 325 and 1075 nm,
with 1 nm of spectral resolution. Hyperspectral signatures from quadrats
serve dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretation. The
workflow developed in this study is presented in
Figure~\ref{fig-workflow}.

@@

The spectral similarities of the reflectance signatures between
intertidal green macrophytes (Magnoliopsida and Chlorophyceae) make
their discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was trained, validated, and applied to
each drone flight (Figure~\ref{fig-workflow}).

@@

The model was applied to all flights at both 12 and 120 m of altitude.
In situ information on georeferenced class type and percent cover
collected during each flight was used to assess the model accuracy,
including. These images were used to construct a validation dataset
indicating the presence or absence of each class. Additionally to the
quadrat-based validation dataset, polygons of each class were photo
interpreted in order to increase the number of pixels of the validation
dataset. A confusion matrix, along with precision metrics such as global
accuracy, sensitivity, specificity, and Kappa coefficient, was generated
for each sites. All validation matrices were then merged to create a
unique matrix . Altogether, a total of 536,000 pixels was used to
validate the model, thus providing a geographically robust validation
dataset.

\subsection{Variable Importance Plots
(VIP)}\label{variable-importance-plots-vip}

\subsection{Impact of the percent cover on the
prediction}\label{impact-of-the-percent-cover-on-the-prediction}

@@

A total of nine prediction maps corresponding to the nine drone flights
were obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the five
vegetation classes on which the model was trained, four were present on
this site, with Chlorophyceae and Rhodophyceae mixed with the seagrass
meadow. There was also Bacillariophyceae forming biofilms on parts of
sediment surface. Although the seagrass bed was solely composed of
\emph{Nanozostera noltei}, various colors of this could be observed from
dark green (corresponding to healthy leaves) to whitish/brown (when
leaves were discolored having lost their pigmentation). Regardless of
color, the meadow was accurately predicted as Magnoliopsida by the
model.

@@

\caption{\label{fig-GafLow}RGB orthomosaic (Left) and Prediction (Right)
of the low altitude flight of Gafanha, Portugal. The total extent of
this flight was 3000m² with a resolution of 8 mm per pixel. Background
colors indicate intertidal area (Light Green) and land area (Light
Grey). The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel
size.}

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
masked in the prediction map. Most of the intertidal area was classified
as Magnoliopsida by the model, including seagrass patches with
discolored leaves. Only a few pixels were classified as Chlorophyceae at
this scale. Furthermore, the area that was classified as
Bacillariophyceae in the low-altitude flight remained mostly classified
as such in the high-altitude flight, though some pixels were classified
as Magnoliopsida. Patches of Rhodophyceae were correctly classified. In
the nothern part of the scene, near the land limit, patches of the
schorre angiosperm \emph{Spartina sp.} were misclassified, either as
Magnoliopsida or as Phaeophyceae.

@@

\caption{\label{fig-GafHigh}RGB orthomosaic (Left) and Prediction
(Right) of the high altitude flight of Gafanha, Portugal. The total
extent of this flight was about 1 km² with a resolution of 80 mm per
pixel. Background colors indicate intertidal area (Light Green), land
area (Light Grey) and water (Light Blue). The red triangle shows the
extent of the low altitude flight of Gafanha presented in
Figure~\ref{fig-GafLow}. The zoom covers an area equivalent to a
10-meter Sentinel-2 pixel size.}

@@

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight was about 1.5 km² with a resolution of
80 mm per pixel. Background colors indicates intertidal area (Light
Green), land area (Light Grey) and water (Light Blue). Each cover an
area equivalent to a 10-meter Sentinel-2 pixel size.}

@@

\caption{\label{fig-Dike}RGB orthomosaic (Top) and Prediction (Bottom)
of the Northern part of Noirmoutier Island, France. The total extent of
this flight was about 28 000 m² with a resolution of 80 mm per pixel.
Background colors indicate intertidal area (Light Green) and land area
(Light Grey). The zoom covers an area equivalent to a 10-meter
Sentinel-2 pixel size.}

@@

A total of 536,000 pixels were used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Kerdrean (5557 pixels) and Marinha Lanzarote (159713 pixels),
respectively. Model global accuracy was 94.26\% with a Kappa coefficient
of 0.92 (Figure~\ref{fig-Validation}). The lowest performing site was
Gafanha High (global accuracy of 75.45\%) whereas Mataducos was the site
with the most accurate prediction (global accuracy of 98.05\%). Overall,
the classes Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae were
correctly classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae was the least performing class (accuracy
of 0.72 ) mainly due to the confusion with Magnoliopsida and Sediment.

@@

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each taxonomic class. The bigger the slice, the more
important the variable for prediction of each class. The right plot
shows the drone raw and standardised reflectance spectra of each class.
Each slice represents the VIP of both raw and standardised reflectance
combined.}

@@

\caption{\label{fig-upscaling}Kernel density plot showing the proportion
of pixel well classified based on the percent cover of the class in high
altitude flight pixels of Gafanha, Portugal. Each subplot shows all the
pixels of the same classes on the high altitude flight. Percent cover of
classes was retrieved using the result of the classification of the low
altitude flight of Gafanha, Portugal. The vertical dashed line shows the
0.85 probability of the model. Everything on the right of this line has
a probability higher then 0.85 and everything on the left of this line
has a lower probability.}
